---
title: "Assignment Wk3"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Info

Creator: Arjun Augustine
Date: January 29, 2021

[Github repository link](https://github.com/ArjunA127/Biol432Assign3)

### Load Packages

```{r message=FALSE}
library(dplyr)
library(knitr)
```

# Wrangle the Fallopia dataset

```{r}
# Load the FallopiaData.csv
dat <- read.csv("InputData/FallopiaData.csv")

# Make a new dataset that only keeps rows in FallopiaData.csv with greater than 60 grams of total biomass
# Only keep the columns 'Total', 'Taxon', 'Scenario', 'Nutrients', in that order from left-right in the new dataset
wran_dat <- dat %>%
  filter(Total >= 60) %>%                   # Remove rows with ‘Total’ biomass < 60
  select(Total, Taxon, Scenario, Nutrients) # Reorder the listed columns, remove all other columns

head(wran_dat)                              # Display first 6 rows of new dataset (with total biomass in milligrams)

# Replace the column "Total" with "TotalG", which converts the ‘Total’ biomass column from mg to grams
wran_dat <- wran_dat %>% 
  transmute(TotalG = Total/1000,            # Make a new column that coverts 'Total' from mg to g
            Total = NULL,                   # Remove original 'Total' column
            Taxon, Scenario, Nutrients)     # Preserve the order of the columns (i.e. keep 'TotalG' where 'Total' was)
# Note: In the transmute function, it was also necessary to re-enter 'Taxon', 'Scenario', and 'Nutrients' to prevent them from being dropped from the dataset

head(wran_dat)                              # Display first 6 rows of new dataset (with total biomass in grams)
```


# Custom function

### Create Custom Function

This function takes two inputs from the user: 1. a vector of data to process (e.g. column from a data.frame object) and 2. a string that defines what calculation to perform ('Average', 'Sum', or 'Observations'). This function ignores missing values when calculating 'Average' or 'Sum', and does not include missing values when counting observed values.

```{r}
cust.fun <- function(x, operator){
  if (operator == "Average"){           # For string 'Average', output the mean of values in the given vector
    out <- mean(x, na.rm = T)
  }
  else if (operator == "Sum"){          # For string 'Sum', output the sum of values in the vector
    out <- sum(x, na.rm = T)
  }
  else if (operator == "Observations"){ # For string 'Observations', output number observed values in the vector
    counter <- 0
    for (i in 1:length(x)){
      if (is.na(x[i]) == F){            # Ensure that any NA is not included when counting observed values
        counter <- counter + 1
      }
    }
    out <- counter
  }
  else{                                 # If any other string is inputted return an error
    print("ERROR. Need to input one of the operations 'Average', 'Sum', or 'Observations'")
  }
  return(out)                           # Output the result for whichever operation the user asked for
}
```

### Use the Custom Function

Use the custom function to count the total number of observations in the 'Taxon' column of the newly wrangled dataset.

```{r}
cust.fun(wran_dat$Taxon, "Observations")
```


Use the custom function to calculate the average TotalG for each of the two Nutrient concentrations.

```{r}
# Calculate average TotalG for the low and high nutrient groups
aTot_nutr <- wran_dat %>% group_by(Nutrients) %>% summarize(Mean = cust.fun(TotalG, "Average"))

# Display a nice looking table for average total mass for each nutrient level
kable(aTot_nutr,
      caption = "Table 1. Average total biomass for different nutrient levels.", # Table caption
      col.names = c("Nutrients", "Mean total mass (g)"),                         # Column headings
      align = 'l')                                                               # Left-align the columns
```
# Export the wrangled datset

Write (i.e. save) the new data to a file called "WrangledData.csv" in the Output folder. 
```{r}
write.csv(wran_dat, "Output/WrangledData.csv")
```

